---
title: "STA 531 HW3"
author: "Daniel Truver"
date: "2/15/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### (1) Rejection Sampling of the Posterior

I doubt this is correct, but I've been staring at it for hours and can't think of anything else.

$$
\begin{aligned}
\pi(\theta\mid y) 
&= \frac{\pi_0(\theta)L(\theta)}{\int_{-\infty}^{\infty}\pi_0(\theta)L(\theta)d\theta}\\
&\leq \frac{\pi_0(\theta)\max_\theta L(\theta)}{\int_{-\infty}^{\infty}\pi_0(\theta)L(\theta)d\theta} \\
&= \frac{\pi_0(\theta)L(\theta_{MLE})}{\int_{-\infty}^{\infty}\pi_0(\theta)L(\theta)d\theta}
\end{aligned}
$$

Then we use rejection sampling on the unnormalized posterior. 

$$
\pi(\theta\mid y) \int_{-\infty}^{\infty}\pi_0(\theta)L(\theta)d\theta \leq \pi_0(\theta)L(\theta_{MLE})
$$

That is, the unnormalized density is bounded above by $c$ times the posterior. Rejection sampling should still work since $\pi_0(\theta)$ is a bonafide probability distribution.

#### (2) Normal-Cauchy Bayes Estimator (MC Integration)

01/31/2018

#### (3) Normal Integration 

Well, there aren't indicator functions mentioned anywhere in my notes, so we proceed by the naive approach of sampling from the uniform distribution, finding the value of function, and then taking the average. For this case, N(0,1), we will sample on the interval [2.5,10]. The variance of our estimate is $(10-2.5)^2/n$ since the variance of the standard normal is 1. So, if we want to be sure of our estimate to 3 digits, we need 
$$
var(\hat{\theta}) = \frac{7.5^2}{n} < 0.0001 \implies n > 562500
$$

```{r mcIntegrationNormal}
set.seed(2018)
n = 562500
a = 2.5; b = 10
points = runif(n, 2.5, 10)
funPoints = dnorm(points, 0, 1)
cat("The estimated P(Z>2.5) is", round((b-a) * mean(funPoints), 4),
    "with the third digit significant.")
```
